---
title: "Modification abundances by amino acid"
author: "COW"
output: html_document
---

Required libraries:
```{r, message=FALSE}
require(dplyr)     # for efficient table manipulations
require(tidyr)     # for efficient table manipulations
require(cowplot)   # for plotting
```

Define a function that can read MODa output matrices:
```{r}
readMODaOutput <- function(sample, moda_folder)
{
  # input filenames
  file_a <- file.path(moda_folder, paste(sample, "a_MODa_PSMs.ptm.txt", sep=''))
  file_b <- file.path(moda_folder, paste(sample, "b_MODa_PSMs.ptm.txt", sep=''))

    # read the raw data
  data_a <- read.table(file_a, header = TRUE, row.names = NULL)
  # R takes the first (unlabeled) column as row names. Convert into numeric column of mass shifts
  colnames(data_a)[1] <- "mass.shift"
  data_a$mass.shift <- as.numeric(data_a$mass.shift)

  # do the same for sample b
  data_b = read.table(file_b, header = TRUE, row.names = NULL)
  colnames(data_b)[1] <- "mass.shift"
  data_b$mass.shift <- as.numeric(data_b$mass.shift)

  # combine the two matrices
  data <- rbind(data_a, data_b)
  # sum entries for the same mass shift
  data <- summarise_each(group_by(data, mass.shift), funs(sum))
  # return data, with column indicating the sample added
  mutate(data, sample = sample)
}
```

Now read in all the data:
```{r}
# read the information about the samples
samples <- read.csv("../DataAnnotations/samples.csv")
# turn to list
sample_list <- as.list(as.character(samples$sample))
names(sample_list) <- samples$sample

data_list <- lapply(sample_list, readMODaOutput, moda_folder = "../MODa_Output/PTMHits/")
# combine all into one table
data_all <- do.call("rbind", data_list)
# combine with other info about samples
data_all <- inner_join(samples, data_all, by = "sample")

head(data_all)
```

# +1 Da mod

We make a bar plot of the percent of modified peptides of a given mass shift, using the variation among samples to estimate errors.

```{r}
# define standard error function
se <- function(x) sqrt(var(x)/length(x))

data_all %>%
  filter(mass.shift==1) %>% # extract the +1 mass shift
  group_by(sample) %>% # analyze by sample
  summarise_each(funs(100*./Sum), -mass.shift) %>% # normalize all and calculate percent
  select(-Sum, -Nterm, -Cterm) -> # remove Sum, Nterm, Cterm columns as we don't plot them
    mass_shifts # store as `mass_shifts`

# now calculate means and standard errors
mass_shifts %>%
  gather(amino.acid, percent, A:Y) %>% # turn table into "long" format
  group_by(amino.acid) %>% # group by amino acid
  summarise(percent.mean = mean(percent), percent.se = se(percent)) ->
   mods_summarized # store as `mods_summarized`

# now plot
plot <- ggplot(mods_summarized, aes(x=amino.acid, y=percent.mean)) +
  geom_bar(stat="identity", fill='red') +
  geom_errorbar(aes(ymin = percent.mean-percent.se, ymax = percent.mean+percent.se)) + 
  ylab("Percent of total modifications") +
  xlab("Modified amino acid")

# save the plot as pdf
#save_plot("plus1_vs_amino_acid.pdf", plot)
plot # display plot
```

# +16 Da mod

We make a bar plot of the percent of modified peptides of a given mass shift, using the variation among samples to estimate errors.

```{r}
data_all %>%
  filter(mass.shift==16) %>% # extract the +16 mass shift
  group_by(sample) %>% # analyze by sample
  summarise_each(funs(100*./Sum), -mass.shift) %>% # normalize all and calculate percent
  select(-Sum, -Nterm, -Cterm) -> # remove Sum, Nterm, Cterm columns as we don't plot them
    mass_shifts # store as `mass_shifts`

# now calculate means and standard errors
mass_shifts %>%
  gather(amino.acid, percent, A:Y) %>% # turn table into "long" format
  group_by(amino.acid) %>% # group by amino acid
  summarise(percent.mean = mean(percent), percent.se = se(percent)) ->
   mods_summarized # store as `mods_summarized`

# now plot
plot <- ggplot(mods_summarized, aes(x=amino.acid, y=percent.mean)) +
  geom_bar(stat="identity", fill='red') +
  geom_errorbar(aes(ymin = percent.mean-percent.se, ymax = percent.mean+percent.se)) + 
  ylab("Percent of total modifications") +
  xlab("Modified amino acid")

# save the plot as pdf
#save_plot("plus1_vs_amino_acid.pdf", plot)
plot # display plot
```

# -2 Da mod

We make a bar plot of the percent of modified peptides of a given mass shift, using the variation among samples to estimate errors.

```{r}
data_all %>%
  filter(mass.shift==-2) %>% # extract the +16 mass shift
  group_by(sample) %>% # analyze by sample
  summarise_each(funs(100*./Sum), -mass.shift, -time, -biol.rep) %>% # normalize all and calculate percent
  select(-Sum, -Nterm, -Cterm) -> # remove Sum, Nterm, Cterm columns as we don't plot them
    mass_shifts # store as `mass_shifts`

# now calculate means and standard errors
mass_shifts %>%
  gather(amino.acid, percent, A:Y) %>% # turn table into "long" format
  group_by(amino.acid) %>% # group by amino acid
  summarise(percent.mean = mean(percent), percent.se = se(percent)) ->
   mods_summarized # store as `mods_summarized`

# now plot
plot <- ggplot(mods_summarized, aes(x=amino.acid, y=percent.mean)) +
  geom_bar(stat="identity", fill='red') +
  geom_errorbar(aes(ymin = percent.mean-percent.se, ymax = percent.mean+percent.se)) + 
  ylab("Percent of total modifications") +
  xlab("Modified amino acid")

# save the plot as pdf
#save_plot("plus1_vs_amino_acid.pdf", plot)
plot # display plot
```

